\Huge Conferences in the Era of Expensive Carbon

### 2019-06-04 main.tex
Some more work on the middle parts

@@

This leads to a puzzle: How do we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  It seems an extremely difficult question to
answer: the harm and the compensating benefits cannot be directly
compared---they are apples and oranges.  
%
Fortunately, there is a well-established mechanism for comparing apples and
oranges (indeed, for comparing almost anything): the magic of markets.  By
mapping choices and preferences in disparate domains to a common scale
(namely money), economic markets can be viewed as massively distributed
computations that balance costs an benefits across vastly different
activities.\bcp{Give a simple, concrete example.}  For this reason, many
policy experts advocate addressing climate change through some form of
carbon pricing, which places a monetary cost on greenhouse gas emissions,
thus giving a market signal that changes aggregate behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}  Exactly how the price of carbon
should be set is a challenging policy question, but it has a clear goal: The
cost should be set so that the world arrives at zero emissions within the
timeframe that climate scientists tell us is safe.

@@

A possible short-term response to both questions involves {\em carbon
  offsets}---a tool by which organizations (and individuals) can voluntarily
impose a price on their own emissions.
%
% What they are:
In a nutshell, the idea is that a conference can raise funds from
registration fees or corporate sponsors and give them, via a carbon offset
{\em vendor}, to a {\em provider} who will use them to support an activity
(e.g., installing methane capture devices on landfill sites or buying
efficient cookstoves to replace open-fire cooking in developing areas) that
avoids emitting an amount of greenhouse gases equivalent to the emissions
from participants' travel to the conference.
%
\begin{itemize}
\item Lots of organizations (universities, companies) are already doing
this; some ACM conferences are starting to experiment with voluntary carbon
offset fees at registration time.
\end{itemize}

\begin{itemize}
\item There are also some significant
concerns.
\item Concern: Organizations are not set up to do it, and the piecemeal
political and bureaucratic work of designing and implementing a separate
policy for each organization is exhausting.
\item Concern: It is hard to choose the price (right now there is an
oversupply), and setting it too low leads to complacency --- it doesn't
really bite, so it doesn't change people's behavior.  The current market
price of carbon is still quite low (around \$15 per ton), whereas the
``social cost of carbon'' is variously estimated at anywhere from \$40 to
\$400 per ton~\bcp{citation}.
Experts agree that the current price will need to rise significantly
over time to achieve the required effect.
\item This issue leads many people to argue that offsets "let people off the
hook" and encourage inaction, but we believe it is consistent with
encouraging action -- especially if offsets are priced right, but even if
(in the short term) they are too cheap -- because it injects a ``price
signal'' that makes emissions something for planners to optimize.
\item We feel that, on balance, voluntary carbon offsets are an effective
tools for raising awareness and changing organizational behavior in the
short term.
\item Moreover, supporting well-chosen carbon mitigation activities has a
clear benefit and can be done now.
\end{itemize}

Beyond imposing a price on our carbon footprint (or waiting for one to be
imposed from outside), a sane response to climate change should clearly also
include taking steps to reduce the size of this footprint.  
\bcp{
  \begin{itemize}
  \item Briefly list a few things we can do
  \begin{itemize}
  \item journalize conferences
  \item go virtual
  \item potential parts of a solution: regional conferences, virtual PC
  meetings, virtual participation, mega-conferences
  \end{itemize}
  \item cite our report
  \end{itemize}
}
%
However, before we undertake large changes, we need to understand the
situation better.  For most of these ideas, it is hard to confidently
predict either the potential benefits (in terms of reduced emissions) or the
potential costs (to scientific exchange, career development of young
scientists, etc.).  Indeed, a preliminary analysis of registration data from
a number of SIGPLAN conferences paints a rather complicated picture.
%
\bcp{Put the discussion of Figure 1 (and an accompanying figure
showing distance traveled to several instances of the same conference) here;
this illustrates the difficulty and the need for careful data gathering and
analysis.} 
All parts of ACM should be doing this kind of analysis.  (The acm-climate
mailing list is one place to exchange ideas about this.)

>>>>

\newcommand{\SECTION}{\paragraph*}

@@

This leads to a puzzle: How do we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  It seems an extremely difficult question to
answer: the harm and the compensating benefits cannot be directly
compared---they are apples and oranges.

\SECTION{Putting a Price on Carbon}

Fortunately, there is a well-established mechanism for comparing apples and
oranges (indeed, for comparing almost anything): the magic of markets.  By
mapping choices and preferences in disparate domains to a common scale
(namely money), economic markets can be viewed as massively distributed
computations that balance costs an benefits across vastly different
activities.\bcp{Give a simple, concrete example.}  For this reason, many
policy experts advocate addressing climate change through some form of
carbon pricing, which places a monetary cost on greenhouse gas emissions,
thus giving a market signal that changes aggregate behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}  Exactly how the price of carbon
should be set is a challenging policy question, but it has a clear goal: The
cost should be set so that the world arrives at zero emissions within the
timeframe that climate scientists tell us is safe.

@@

\SECTION{Carbon Offsets}

In the short term, one possible response to both questions is to use {\em
  carbon offsets}, which allow organizations (and individuals) to
voluntarily impose a price on their own emissions~\cite{CarbonOFfsetReport}.
%
In a nutshell, the idea is that a conference can raise funds from
registration fees or corporate sponsors and give them, via a carbon offset
{\em vendor}, to a {\em provider} who will use them to support an activity
(e.g., installing methane capture devices on landfill sites or buying
efficient cookstoves to replace open-fire cooking in developing areas) that
is {\em certified} (by yet another organization) to permanently avoid
emitting an amount of greenhouse gases equivalent to the emissions from
participants' travel to the conference.
%
Many organizations (including companies such as Google, Dell, Microsoft,
General Motors, Delta Airlines, Lyft, and Expedia, numerous universities,
and a number of academic societies) are already buying carbon offsets, and
some ACM conferences have already started to experiment with voluntary
carbon offset fees at registration time.

Carbon offsets do come with some significant concerns.
%
A superficial but potentially sigificant issue is that most
organizations do not currently have accounting procedures in place to deal
with them, and the piecemeal political and bureaucratic work of designing
and implementing a separate policy for each organization can consume
significant energy.
%
A more fundamental problem is that it is not straightforward to decide how
much it should cost to offset a ton of carbon: current prices vary widely
around an average of around \$15 per ton, depending on the vendor, the
nature of the offset project, the region of the world where it will be
implemented, etc..  Worse, in order for offsets (or any other form of a
carbon pricing) to fully motivate the needed changes in behavior, the price
may need to go much higher---the true ``social cost'' of a ton of carbon is
variously estimated at anywhere from \$40 to \$400 per
ton~\bcp{citation}---and setting it too low may encourage complacency.
%
And finally, 
%
We feel that, on balance, voluntary carbon offsetting can be an
effective tool for raising awareness and changing organizational behavior
in the short term.
%
Moreover, supporting well-chosen carbon mitigation activities has a clear
benefit in reducing global net emissions now.

\SECTION{Measure, Analyze, Optimize}

Beyond imposing a price on our carbon footprint (or waiting for one
to be imposed from outside), a sane response to climate change should
clearly also include taking steps to reduce the size of this footprint.  For
conferences, there are many options that one might
contemplate~\cite{ClimateCommitteReport}, from easy 
but incremental changes like making program committee meetings virtual
instead of physical and enabling virtual participation via live-streaming,
all the way to complete re-imagining of conferences and their role in
science (holding conferences completely in virtual spaces, replacing
conferences with journals, etc.), with some intriguing hybrid forms in the
middle (regional conferences, mega-conferences, conferences split into
two locations on different continents, etc.).  
%
However, before we undertake large changes, we need to understand the
situation better.  For most of these ideas, it is hard to confidently
predict either the potential benefits (in terms of reduced emissions) or the
potential costs (to scientific exchange, career development of young
scientists, etc.).  Indeed, a preliminary analysis of registration data from
a number of SIGPLAN conferences paints a rather complicated picture.
%
\bcp{Put the discussion of Figure 1 (and an accompanying figure
showing distance traveled to several instances of the same conference) here;
this illustrates the difficulty and the need for careful data gathering and
analysis.} 
All parts of ACM should be doing this kind of analysis.  (The acm-climate
mailing list is one place to exchange ideas about this.)

\SECTION{Deeper Changes}
